---
title: "PathQuant vignette"
author: "Sandra Therrien-Laperrière"
date: "4 avril 2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

##Description  

PathQuant aims to help quantify associations between Genomic and 
Metabolomic Data. To do so, PathQuanr calculates distance in terms od reactions 
between Genes and Metabolites mapped on KEGG's metabolic pathway maps.



## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.



    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Fonctions

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.width=7, fig.height=4}
library("PathQuant")

heatmapFunction(metabolismOverviewMapKEGGId, shinAndAlDF)

```


```{r, fig.width=7,fig.height=4 }

distanceGeneToAllMetabolite("hsa01100", shinAndAlDF, completeMetaboDF, "hsa:1373")
distanceGeneToAllMetabolite("hsa01100", shinAndAlDF, completeMetaboDF, "hsa:27165")

```
```{r, size=12}

ASP <- getDistanceAsso("hsa01100", shinAndAlDF, F)
ASP[1:10,]

ASP1 <- getDistanceAll("hsa01100", shinAndAlDF, completeMetaboDF)
ASP1[1:10,]
```











